A Plasma Torus Around a Young Low-Mass Star

### 2025-03-24 paper/ms.tex
update

\newcommand{\starname}{TIC 141146667}

@@

%%%%%%%%%%%%%%%%
% INSTITUTIONS %
%%%%%%%%%%%%%%%%
\newcommand{\carnegie}{Observatories of the Carnegie Institution for Science, Pasadena, CA 91101, USA}
%%%%%%%%%%%%%%%%

@@

\scriptsize
\begin{affiliations}
\item \carnegie
\item Carnegie Fellow
\end{affiliations}
\normalsize

@@

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/f2.pdf}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 2 (Movie):}  
  Emission from circumstellar plasma orbiting TIC 141146667.
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{available here},
  shows the spectral evolution over five hours.
  {\bf a,} Average TESS light curve from 5 February 2024 to 26
  February 2024 (UT) folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra acquired on 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for $v_{\rm eq}$=130\,\kms).  While
  emission in the line core originates in the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting the time-averaged
  spectrum.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch
  is set to match the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure}

@@

% ** keck/hires: emission beyond v/veq>1
The spectroscopy shows emission beyond the star's equatorial velocity
($v_{\rm eq}$=130\,\kms).  There are at least two distinct emission
components, separated by half a cycle in phase.  The first component
has clearer sinusoidal behaviour and is double-peaked, with peak
semi-amplitudes of $K_1$=2.1\,$v_{\rm eq}$ and 2.7\,$v_{\rm eq}$.
There is significant non-periodic variability in the emissivity of
this double-peaked component: the flux excess begins with an amplitude
of 70\% of the continuum flux early in the observation sequence, and
diminishes to 10\% by its end.  The component 180$^\circ$ opposite in
phase is detected only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this
latter component appears connected to the star in velocity space.
While its peak semi-amplitude of $K_1$=3.9\,$v_{\rm eq}$ is achieved
at both $\phi$=0.25 and 0.75, its amplitude similarly decreases from a
60\% excess over the continuum at the beginning of the observation
sequence to a 10\% excess by its end.  The sinusoidal period for all
of these emission components is consistent with the photometric
\periodhr\ hour period.

@@

% * keck/hires: behavior at v/veq<1
Within the stellar H$\alpha$ line core, at $|\Delta v /
v_{\rm eq}|<1$, we observe a superposition of emission from the star's
chromosphere,
modulated by any occulting material capable of absorbing or
emitting H$\alpha$ photons.  In Figure~\ref{fig:spec}e, the data from
$\phi$=0.4-1.2 have a simple interpretation: from $\phi$=0.4-0.9, a hot
(emitting) region first gradually crosses the stellar line profile, followed from
$\phi$=0.7-1.2 by the transit of a cool (absorbing) region. Phases $\phi$$<$0.4
show a mix of similar events, though the time sampling is
sufficiently coarse that the interpretation is less clear.  A final
exercise to quantify the behavior in the line core is shown in
Figure~\ref{fig:spec}d, where $f_{\rm H\alpha\ core}$ denotes the
summed flux at $|\Delta v / v_{\rm eq}|<1$.  This panel shows that
changes in the line core flux are correlated with the
broadband variability over most of the light curve, except at
$\phi$$\approx$0.5, during the transit of the 3.9\,$R_\star$ clump and
the occultation of the lower-velocity clump.

@@

Spectra of magetically-active, rapidly rotating stars with a wide range
of masses have been known to exhibit both sinusoidal emission features
\cite{Donati2000,Townsend2005,Dunstone2006,Skelly2008} as well as
transient absorption features
\cite{CollierCameron1989,CollierCameron1992,Cang2020} similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \cite{Bouma2024}.   One interpretation for
such spectroscopic variability comes from an analogy to quiescent solar
prominences, which are cool condensations of plasma in the solar corona
that can last days to weeks \cite{VialEngvold2015}.  These
condensations fall back to the solar surface because gravity is
stronger than any magnetic tension or centrifugal force capable of
sustaining them.  However for stars with magnetospheric radii $R_{\rm
m}$ that exceed their corotation radii $R_{\rm c}$, the effective
potential experienced by a charged particle can have a local minimum
outside $R_{\rm c}$, enabling the material to accumulate over much
longer timescales \cite{Petit2013,Daley-Yates2024}.  Generally
speaking, such material need neither transit, nor be optically thick.
For instance, the H$\alpha$ emission clump at 3.9\,$R_\star$ shows no 
photometric signal in the February 2024 photometry.

Our Keck/HIRES observations are the first reported time-series
spectra of a CPV, and they show that corotating circumstellar plasma clumps 
are present around at least one such star.
This
observation rules out a ``starspot-only'' origin scenario for CPVs,
\cite{Koen2021} because such scenarios have no means of explaining
spectroscopic emission beyond the stellar disk.  Similarly, scenarios in
which the circumstellar material is made only of dust are also ruled
out.  While dust may be present, to explain the H$\alpha$ emission the
clumps must include plasma with a significant population of hydrogen
atoms in the $n$=3 excited state.  
More quantitatively, the observed H$\alpha$ luminosity suggests
characteristic densities and masses for the gaseous component of these
clumps of $n_{\rm H} \sim 10^{11}$-$10^{12}$\,cm$^{-3}$ and $M_{\rm
gas} \sim 10^{17}$\,g (see Section~\ref{subsec:gas}).
Dust, if present, is independently constrained to have a total
mass $10^{15}\,{\rm g} < M_{\rm dust} < 10^{17}\,{\rm g}$ (see
Section~\ref{subsec:dust}).
This circumstellar material could plausibly originate from
three sites: the star, an old and undetected disk, or outgassing rocky
bodies.  This latter possibility would render CPVs as extrasolar analogs
of the Jupiter-Io plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars with radiatively-driven winds that
accumulate into warped plasma tori \cite{Townsend2005,Townsend2008}.
These tori tend to have dense antipodal accumulations of plasma sculpted
by tilted-dipole magnetic fields, and the transits of these clumps are
thought to produce the observed broadband photometric variability
through bound-free scattering \cite{Townsend2005} and Thomson scattering
\cite{Berry2022}.  For $\sigma$~Ori~E and almost all of its analogs, the
result is light curves that appear ``simple'', resembling those of
eclipsing binaries, and time-dynamic H$\alpha$ spectra similar to those
in Figure~\ref{fig:spec} \cite{Townsend2005,Townsend2008}.  The two known exceptions,
HD~37776 and HD~64740, show complex light curves resembling CPVs
\cite{Mikulasek2020,Bouma2024} and have spectropolarimetric magnetic
field maps indicating strong contributions from higher-order magnetic
moments \cite{Kochukhov2011,Shultz2018}.  There are two implications:
first, the complexity of CPVs may be a direct consequence of magnetic
fields with highly multipolar contributions.  Second, CPVs can be a
source of astrophysical false positives in photometric classification
of eclipsing binaries and transiting exoplanets around young pre-main
sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The material's composition -- either pure plasma, or else a dusty plasma
-- can be clarified by time-series optical and infrared
spectrophotometry.  While observations of CPVs in the optical have
previously suggested chromaticity consistent with dust
\cite{Tanimoto2020,Gunther2022,Koen2023}, a gray opacity source such as
electron scattering in a plasma transiting over starspots might also
produce chromatic features \cite{Rackham2018}.  The composition and size
distribution of any dust that is present could be determined
by measuring the extinction curve for one or more CPVs from
1-10\,$\mu$m.  Composition and size distributions similar to debris from
rocky bodies seen around white dwarfs \cite{Reach2009} would indicate a
rocky-body origin.  Compositions and sizes similar to the interstellar
medium could be indicative of condensed dust in an M dwarf wind,
similar to that formed in the environments of more evolved stars
\cite{Marigo2008}.

The role of the star's magnetic field could be better understood through
new observations, and new theory.  From the theoretical perspective,
there is an urgent need for rigid-field magneto-hydrodynamic modeling to
go beyond previous work \cite{Townsend2005,Townsend2008,Krticka2022} and
to explore what field topologies might explain the observed light curve
morphologies \cite{Bouma2024}.  In particular, dynamo simulations of
fully-convective M dwarfs have suggested that global-scale mean fields
might be confined to a single hemisphere \cite{Brown2020}; such fields
could yield accumulation surfaces quite different from those that have
previously been explored.  Observationally, spectropolarimetry has the
potential to assess both the field strength and topology.  An
independent probe could also be connected the recent work
\cite{Kaur2024} showing that CPVs are variable radio emitters with
emission components that can be not only persistent, but also
short-lived and highly polarized.  In particular, detecting radio
emission produced by an electron cyclotron maser instability (e.g.
\cite{Callingham2021}) would provide a measurement of the field strength
at the site of the emitting region, further clarifying the magnetic
field's structure.

@@

{\it No Evidence For Binarity}---Any significant periodicity in the
radial velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This sets
an upper limit on the mass of any putative companion at the four hour
period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE), a proxy for the goodness of fit for a single-source
astrometric model to the Gaia astrometry, is 1.23, within the usual
range for apparently single sources.  There are no resolved sources in
the Gaia DR3 point source catalog.  Finally, we checked the TESS light
curve for evidence of secondary photometric periods by subtracting the
mean CPV signal over each sector and performing a phase-dispersion
minimization analysis \cite{Stellingwerf1978,2021zndo...1011188B}.
There were no secondary periods in the TESS data.  Previous work
\cite{Bouma2024} has shown that about 30\% of CPVs show evidence for
excess noise above the Gaia single-source astrometric model, and about
40\% of CPVs show evidence for unresolved binary companions based on the
presence of secondary photometric periods.  This agrees with analyses
showing that multi-periodic low-mass stars are generally unresolved
binaries \cite{Tokovinin2018}.  Overall, the CPV binary fraction seems
consistent with that for field M dwarfs \cite{Winters2019}, pointing to
a weak or non-existent connection between the CPV phenomenon and
(wide) binarity.  For TIC~141146667 specifically, while we find no evidence for
stellar multiplicity, our data are minimally constraining for the
scenario of a low-luminosity companion ($F_1$/$F_2$$\lesssim$0.1) at
with apparent separations below 0$\farcs$1.

@@

{\it Effective temperature, radius, mass, and spectral classification}---We adopted
the star's effective temperature and radius measured using the spectral
energy distribution (SED) fitting procedure described by
\cite{Bouma2024}.  To summarize, this approach used
\texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we performed the fit using the broadband magnitudes from
Gaia DR2, APASS, 2MASS, SDSS, and WISE $W1$ and $W2$.  The resulting
best-fit SED is shown in Figure~\ref{fig:sed}.  This method has the most
constraining power for the star's effective temperature (2972 $\pm$
40\,K) and radius (0.42 $\pm$ 0.02\,$R_\odot$); the surface gravity,
metallicity, and reddening are only weakly constrained.  We
determined the star's spectral type to be M5.5Ve by visually comparing
the HIRES spectra against the photometric standards tabulated by
\cite{Bochanski2007}.   This result agrees with the effective
temperature found from the SED fitting \cite{Pecaut2013}.
We measured the equivalent width of the H$\alpha$ line by fitting a
range of models to the time-averaged line profile shown in
Figure~\ref{fig:spec}, selecting the model that minimized the Bayesian
information criterion (we found a sum of two Gaussians to be preferred), and numerically
integrating this best fit model.
Our quoted result, EW$_{{\rm H}\alpha}$=7.2$\pm$0.2\,\AA, comes from
numerically integrating within $|\Delta v/v_{\rm eq}|<1$; 
integrating over the entire line profile (including the broad wings) would yield
EW$_{{\rm H}\alpha}$=10.2$\pm$0.3\,\AA
Either value would classify the star as a weak-lined T Tauri \cite{Briceno2019}.

@@

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found that over 90\% of CPVs are associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is in the minority.  We calculated the probability of
TIC~141146667 being part of any of the nearby known groups using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  That particular model
classifies it as a field star at $>$99.9\% confidence.  We also searched
the local vicinity of TIC~141146667 for neighbors with similar projected
on-sky velocities using \texttt{comove} \cite{Tofflemire2021}.  This
yielded no strong candidates for co-moving stars with projected
tangential velocities $\Delta v_{\rm T}$$<$5\,\kms\ that share its
isochronal youth.

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main sequence M dwarf, similar to
all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.  The
star's proximity ($d$=58\,pc) and its high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.  Figure~\ref{fig:camd} shows the
location of the star in the color--absolute magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
IC~2602, and the Pleiades.  To make this diagram, we adopted the USco
members in the $\delta$~Sco and $\sigma$~Sco sub-associations from
\cite{Ratzenbock2023}, and the IC~2602 and Pleiades members from
\cite{Hunt2024}.  We assumed an average $V$-band extinction $A_{\rm
V}$=$\{0.12, 0.11, 0.10\}\,{\rm mag}$ for USco \cite{Pecaut2016},
IC~2602 \cite{Hunt2024}, and the Pleiades \cite{Hunt2024} respectively.
We dereddened the photometry using the extinction coefficients
$k_X\equiv A_X/A_0$ tabulated in \cite{GaiaCollaboration2018}, assuming
that $A_0 = 3.1 E(B-V)$.

@@

{\it Age: Lithium}---The depletion of lithium due to $^7$Li(p,
$\alpha$)$^4$He burning in the cores of low-mass stars has been studied
for over sixty years \cite{Hayashi1963,Bildsten1997,Burke2004}.
\cite{Wood2023} provided an instructive recent overview: an abbreviated
summary is that sufficiently cool and young M dwarfs show the 6707.8\,\AA\
doublet in absorption, $\gtrsim$10\% below their ``continua''.  Unlike
for Sun-like stars, the continuum for M dwarfs is poorly defined due to
their molecular absorption.  We attempted a lithium measurement by
constructing a wavelength-binned and Doppler-corrected TIC~141146667
spectrum, assigning its uncertainties based on the measured scatter
across the five hour dataset.  We then compared this average spectrum
against the nearest matching template from \cite{Bochanski2007}.  The
data show a small depression near the expected lithium wavelength,
potentially consistent with the $\Delta \lambda$$\approx$3\,\AA\ line
broadening.  This feature nominally yields ${\rm EW}_{\rm
Li}$=71$^{+18}_{-13}$\,m\AA, where the statistical uncertainties are
evaluated using a bootstrap resampling technique from the statistical
uncertainties in the spectrum.  However the systematic uncertainties associated
with the continuum normalization are larger, and likely comparable to
the amplitude of the feature; we therefore treat the result of this
measurement as a $2\sigma$ upper limit: ${\rm EW}_{\rm Li}$$<$107\,m\AA.

@@

{\it Age: Summary}---The main indicators for the youth of
TIC~141146667 are {\it i)} that it is a complex periodic variable, and
{\it ii)} that it is 1.5 magnitudes brighter than main sequence stars
of the same color, while showing no indicators for binarity.  The
first item constitutes evidence for youth since all known CPVs are
$\lesssim$200\,Myr \cite{Bouma2024}.  The reference sequences for the
Pleiades (112\,Myr, \cite{Dahm2015}) show a few stars of equal
luminosity and the same temperature, suggesting a photometric
isochronal age upper limit $\lesssim$150\,Myr.  The weak lithium
absorption suggests an age of at least 20\,Myr based on an empirical
comparison using Gaia-ESO spectra, or at least 35\,Myr based on the
\cite{Feiden2016} isochrones.  These considerations yield our adopted
age range of 150\,Myr.

>>>>

@@

%%%%%%%%%%%%%%%%
% INSTITUTIONS %
%%%%%%%%%%%%%%%%
\newcommand{\carnegie}{Observatories of the Carnegie Institution for Science, Pasadena, CA 91101, USA}
\newcommand{\caltech}{Department of Astronomy, California Institute of Technology, Pasadena, CA 91125, USA}
%%%%%%%%%%%%%%%%

@@

\scriptsize
\begin{affiliations}
\item \carnegie
\item \caltech
\item Carnegie Fellow; 51 Pegasi b Fellow
\end{affiliations}
\normalsize

@@

\begin{figure}[!tp]
  \centering
  \includegraphics[width=0.99\textwidth]{figures/f2.pdf}
  \captionsetup{labelformat=empty}
  \caption{{\bf Figure 2 (Movie):}  
  Emission from circumstellar plasma orbiting TIC 141146667.
  The online movie,
  \href{https://lgbouma.com/movies/TIC141146667_sixpanel.mp4}{available here},
  shows the spectral evolution over five hours.
  {\bf a,} Average TESS light curve from 5 February 2024 to 26
  February 2024 folded on the \periodhr\ hour period.  Black
  points are phase-averaged; gray are the raw data.
  {\bf b,} Keck/HIRES H$\alpha$ spectra acquired on 17
  February 2024.  The continuum is set to unity, and the darkest color
  is at twice the continuum to accentuate emission outside the line
  core ($|v/v_{\rm eq}|>1$, for $v_{\rm eq}$=130\,\kms).  While
  emission in the line core originates in the star's chromosphere,
  the sinusoidal emission features are most readily described by a
  warped plasma torus.
  {\bf c,} Individual epochs of Panel b, visible in the
  online movie.  The dotted line shows the time-averaged spectrum,
  $f_{\langle t \rangle}$.
  {\bf d,} As in Panel a, but overplotting the
  median-normalized H$\alpha$ light curve at $|v/v_{\rm eq}|<1$.
  {\bf e,} As in Panel b, after subtracting the time-averaged
  spectrum.  The line core shows H$\alpha$ excesses and decrements
  advancing from the blue to red wings.
  The asymmetric color stretch
  is set to mirror the dynamic range of the data.
  {\bf f,} Individual epochs of Panel e, visible in the online
  movie.}
  \label{fig:spec}
\end{figure}

@@

% ** keck/hires: emission beyond v/veq>1
The spectroscopy shows emission beyond the star's equatorial velocity
($v_{\rm eq}$=130\,\kms).  There are at least two distinct emission
components, separated by half a cycle in phase.  The first component
has clearer sinusoidal behaviour and is double-peaked, with peak
semi-amplitudes of $K_1$=2.1\,$v_{\rm eq}$ and 2.7\,$v_{\rm eq}$.
There is significant non-periodic variability in the emissivity of
this double-peaked component: the flux excess begins with an amplitude
of 70\% of the continuum flux early in the observation sequence, and
diminishes to 10\% by its end.  The component 180$^\circ$ opposite in
phase is detected only from $\phi$=0.2-1.0.  From $\phi$=0.2-0.5, this
latter component appears connected to the star in velocity space.
While its peak semi-amplitude of $K_1$=3.9\,$v_{\rm eq}$ is achieved
at both $\phi$=0.25 and 0.75, its amplitude similarly decreases from a
60\% excess over the continuum at the beginning of the observation
sequence to a 10\% excess by its end.  The apparent period for all
of these emission components is consistent with the photometric
\periodhr\ hour period.

@@

% * keck/hires: behavior at v/veq<1
The stellar H$\alpha$ line core, at $|\Delta v / v_{\rm
eq}|<1$, captures a superposition of emission from the star's
chromosphere, modulated by any occulting material capable of absorbing
or emitting H$\alpha$ photons.  In Figure~\ref{fig:spec}e, the data
from $\phi$=0.4-1.2 have a simple interpretation: from $\phi$=0.4-0.9,
a hot (emitting) region first gradually crosses the stellar line
profile, followed from $\phi$=0.7-1.2 by the transit of a cool
(absorbing) region. Phases $\phi$$<$0.4 show a mix of similar events,
though the time sampling is sufficiently coarse that the
interpretation is less clear.  A final exercise to quantify the
behavior in the line core is shown in Figure~\ref{fig:spec}d, where
$f_{\rm H\alpha\ core}$ denotes the summed flux at $|\Delta v / v_{\rm
eq}|<1$.  This panel shows that changes in the line core flux are
correlated with the broadband variability over most of the light
curve, except at $\phi$$\approx$0.5, during the transit of the
3.9\,$R_\star$ clump and the occultation of the lower-velocity clump.

@@

Spectra of magetically-active, rapidly rotating stars with a wide range
of masses have been known to exhibit both sinusoidal emission features
\cite{Donati2000,Townsend2005,Dunstone2006,Skelly2008} and
transient absorption features
\cite{CollierCameron1989,CollierCameron1992,Cang2020} similar to
Figure~\ref{fig:spec}.  No such stars were previously known to show
complex light curves \cite{Bouma2024}.   One interpretation for
such spectroscopic variability comes from an analogy to quiescent solar
prominences, which are cool condensations of plasma in the solar corona
that can last days to weeks \cite{VialEngvold2015}.  These
condensations fall back to the Sun's surface because gravity is
stronger than any magnetic tension or centrifugal force capable of
sustaining them.  However for stars with magnetospheric radii $R_{\rm
m}$ that exceed their corotation radii $R_{\rm c}$, the effective
potential experienced by a charged particle can have a local minimum
outside $R_{\rm c}$, enabling the material to accumulate over much
longer timescales \cite{Petit2013,Daley-Yates2024}.  Generally
speaking, such material need neither transit, nor be optically thick.
For instance, the H$\alpha$ emission clump at 3.9\,$R_\star$ shows no 
photometric signal in the February 2024 photometry.

Our Keck/HIRES observations are the first reported time-series spectra
of a CPV, and they show that corotating circumstellar plasma clumps
are present around at least one such star.  This observation rules out
a ``starspot-only'' origin scenario for CPVs, \cite{Koen2021} because
such scenarios have no means of explaining spectroscopic emission
beyond the stellar disk.  Similarly, scenarios in which the
circumstellar material is made only of dust are also ruled out.  While
dust may be present, to explain the H$\alpha$ emission the clumps must
include plasma with a significant population of hydrogen atoms in the
$n$=3 excited state, for $n$ the principal quantum number.  More quantitatively, the observed H$\alpha$
luminosity suggests characteristic densities and masses for the
gaseous component of these clumps of $n_{\rm H} \sim
10^{11}$-$10^{12}$\,cm$^{-3}$ and $M_{\rm gas} \sim 10^{17}$\,g (see
Section~\ref{subsec:gas}).  Dust, if present, is independently
constrained to have a total mass $10^{15}\,{\rm g} < M_{\rm dust} <
10^{17}\,{\rm g}$ (see Section~\ref{subsec:dust}).  This circumstellar
material could originate either from the star or from an external
source.  Plausible external sources include an old and undetected
disk, comets, or a close-in exoplanet.
This latter possibility would render CPVs as extrasolar analogs of the
Jupiter-Io plasma torus (e.g.~Ref.~\cite{Bagenal1981}).

The other potential analog for the CPVs are the $\sigma$~Ori~E
variables, a rare subset of B stars with radiatively-driven winds that
accumulate into warped plasma tori \cite{Townsend2005,Townsend2008}.
These tori tend to have dense antipodal accumulations of plasma
sculpted by tilted-dipole magnetic fields, and the transits of these
clumps produce the observed broadband photometric
variability through bound-free scattering \cite{Townsend2005} and
Thomson scattering \cite{Berry2022}.  For $\sigma$~Ori~E and almost
all of its analogs, the result is light curves that appear ``simple'',
resembling those of eclipsing binaries, and time-dynamic H$\alpha$
spectra similar to those in Figure~\ref{fig:spec}
\cite{Townsend2005,Townsend2008}.  The two known exceptions, HD~37776
and HD~64740, show complex light curves resembling CPVs
\cite{Mikulasek2020,Bouma2024} and have spectropolarimetric magnetic
field maps indicating strong contributions from higher-order magnetic
moments \cite{Kochukhov2011,Shultz2018}.  There are two implications:
first, the complexity of CPVs may be a direct consequence of magnetic
fields with highly multipolar contributions.  Second, CPVs can be a
source of astrophysical false positives in photometric classification
of eclipsing binaries and transiting exoplanets around young pre-main
sequence M dwarfs \cite{Johns-Krull2016,Bouma2020}.

@@

The material's composition -- either pure plasma, or else a dusty
plasma -- can be clarified by time-series optical and infrared
spectrophotometry.  While observations of CPVs in the optical may
suggest chromaticity consistent with dust
\cite{Tanimoto2020,Gunther2022,Koen2023}, a gray opacity source such
as electron scattering in a plasma transiting over starspots might
also produce chromatic features \cite{Rackham2018}.  The composition
and size distribution of any dust that is present could be determined
by measuring the extinction curve for one or more CPVs from
1-10\,$\mu$m.  Composition and size distributions similar to debris
from rocky bodies seen around white dwarfs \cite{Reach2009} would
indicate a rocky-body origin.  Compositions and sizes similar to the
interstellar medium could be indicative of condensed dust in an M
dwarf wind, similar to that formed in the environments of more evolved
stars \cite{Marigo2008}.

The role of the star's magnetic field could be better understood
through new observations, and new theory.  From the theoretical
perspective, there is an urgent need for rigid-field
magneto-hydrodynamic modeling to go beyond previous work
\cite{Townsend2005,Townsend2008,Krticka2022} and to explore what field
topologies might explain the observed light curve morphologies
\cite{Bouma2024}.  In particular, dynamo simulations of
fully-convective M dwarfs have suggested that global-scale mean fields
might be confined to a single hemisphere \cite{Brown2020}; such fields
would yield accumulation surfaces quite different from those that have
previously been explored.  Observationally, spectropolarimetry has the
potential to assess both the field strength and topology.  An
independent probe could also be connected the recent work
\cite{Kaur2024} showing that CPVs are variable radio emitters with
emission components that can be not only persistent, but also
short-lived and highly polarized.  In particular, detecting radio
emission produced by an electron cyclotron maser instability (e.g.
\cite{Callingham2021}) would provide a measurement of the field
strength at the site of the emitting region, further clarifying the
magnetic field's structure.

@@

{\it No Evidence For Binarity}---Any periodicity in the
radial velocity time-series is ruled out at the rotation period for
semi-amplitudes above 2.85\,\kms\ (at 3$\sigma$ confidence).  This
sets an upper limit on the mass of any putative companion at the four
hour period of $m \sin i $$<$2.4\,$M_{\rm Jup}$.  Regarding possible
companions at wider separations, the Gaia DR3 renormalized unit weight
error (RUWE), a proxy for the goodness of fit for a single-source
astrometric model to the Gaia astrometry, is 1.23, within the usual
range for apparently single sources.  There are no resolved companions in
the Gaia DR3 point source catalog.  Finally, we checked the TESS light
curve for evidence of secondary photometric periods by subtracting the
mean CPV signal over each sector and performing a phase-dispersion
minimization analysis \cite{Stellingwerf1978,2021zndo...1011188B}.
There were no secondary periods in the TESS data.  Previous work
\cite{Bouma2024} has shown that about 30\% of CPVs show evidence for
excess noise above the Gaia single-source astrometric model, and about
40\% of CPVs show evidence for unresolved binary companions based on
the presence of secondary photometric periods.  This agrees with
analyses showing that multi-periodic low-mass stars are generally
unresolved binaries \cite{Tokovinin2018}.  Overall, the CPV binary
fraction seems consistent with that for field M dwarfs
\cite{Winters2019}, pointing to a weak or non-existent connection
between the CPV phenomenon and (wide) binarity.  For TIC~141146667
specifically, although we find no evidence for stellar multiplicity, our
data are minimally constraining for the scenario of a low-luminosity
companion ($F_1$/$F_2$$\lesssim$0.1) with apparent separation
below 0$\farcs$1.

@@

{\it Effective temperature, radius, mass, and spectral
classification}---We adopted the star's effective temperature and
radius measured using the spectral energy distribution (SED) fitting
procedure described by \cite{Bouma2024}.  To summarize, this approach
used \texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we performed the fit using the broadband magnitudes from
Gaia DR2, APASS, 2MASS, SDSS, and WISE $W1$ and $W2$.  The resulting
best-fit SED is shown in Figure~\ref{fig:sed}.  This method has the
most constraining power for the star's effective temperature (2972
$\pm$ 40\,K) and radius (0.42 $\pm$ 0.02\,$R_\odot$); the surface
gravity, metallicity, and reddening are only weakly constrained.  We
determined the star's spectral type to be M5.5Ve by visually comparing
the HIRES spectra against the photometric standards tabulated by
\cite{Bochanski2007}.   This result agrees with the effective
temperature found from the SED fitting \cite{Pecaut2013}.  We measured
the equivalent width of the H$\alpha$ line by fitting a range of
models to the time-averaged line profile shown in
Figure~\ref{fig:spec}, selecting the model that minimized the Bayesian
information criterion, and numerically integrating this best fit model.  
We found a sum of two Gaussians to be the preferred model; our quoted
result, EW$_{{\rm H}\alpha}$=7.2$\pm$0.2\,\AA, comes from numerically
integrating within $|\Delta v/v_{\rm eq}|<1$.  Integrating over the
entire line profile, including the broad wings, would yield EW$_{{\rm
H}\alpha}$=10.2$\pm$0.3\,\AA. Either value would classify the star as
a weak-lined T Tauri \cite{Briceno2019}.

@@

{\it Age: No Obvious Association Membership}---Previous work
\cite{Bouma2024} has found that over 90\% of CPVs are associated with
known young moving groups based on their positions and kinematics.
TIC~141146667 is in the minority.  We calculated the probability of
TIC~141146667 being part of any nearby known group using
BANYAN\,$\Sigma$ v1.2 \cite{Gagne2018}.  That particular model
classifies it as a field star at $>$99.9\% confidence.  We also
searched the local vicinity of TIC~141146667 for neighbors with
similar projected on-sky velocities using \texttt{comove}
\cite{Tofflemire2021}.  This yielded no strong candidates for
co-moving stars with projected tangential velocities $\Delta v_{\rm
T}$$<$5\,\kms\ that share its isochronal youth.

{\it Age: Isochrones}---The color and absolute magnitude of
TIC~141146667 suggest that it is a pre-main sequence M dwarf, similar to
all other known CPVs \cite{Stauffer2017,Stauffer2021,Bouma2024}.  The
star's proximity ($d$=58\,pc) and high galactic latitude
($b$=$+$53$^\circ$) yield negligible interstellar reddening along the
line of sight \cite{Green2019}.  Figure~\ref{fig:camd} shows the
location of the star in the color--absolute magnitude diagram (CAMD)
relative to young stellar populations including Upper Scorpius (USco),
IC~2602, and the Pleiades.  To make this diagram, we adopted the USco
members in the $\delta$~Sco and $\sigma$~Sco sub-associations from
\cite{Ratzenbock2023}, and the IC~2602 and Pleiades members from
\cite{Hunt2024}.  We assumed an average $V$-band extinction $A_{\rm
V}$=$\{0.12, 0.11, 0.10\}\,{\rm mag}$ for USco \cite{Pecaut2016},
IC~2602 \cite{Hunt2024}, and the Pleiades \cite{Hunt2024} respectively.
We dereddened the photometry using the extinction coefficients
$k_X\equiv A_X/A_0$ tabulated in \cite{GaiaCollaboration2018}, assuming
that $A_0 = 3.1 E(B-V)$.

@@

{\it Age: Lithium}---The depletion of lithium due to $^7$Li(p,
$\alpha$)$^4$He burning in the cores of low-mass stars has been studied
for over sixty years \cite{Hayashi1963,Bildsten1997,Burke2004}.
\cite{Wood2023} provided a recent overview: an abbreviated
summary is that sufficiently cool and young M dwarfs show the 6707.8\,\AA\
doublet in absorption, $\gtrsim$10\% below their ``continua''.  Unlike
for Sun-like stars, the continuum for M dwarfs is poorly defined due to
their molecular absorption.  We therefore attempted a lithium measurement by
constructing a wavelength-binned and Doppler-corrected TIC~141146667
spectrum, assigning its uncertainties based on the measured scatter
across the five hour dataset.  We then compared this average spectrum
against the nearest matching M6 template from \cite{Bochanski2007}.  The
data show a small depression near the expected lithium wavelength,
potentially consistent with the $\Delta \lambda$$\approx$3\,\AA\ line
broadening.  This feature nominally yields ${\rm EW}_{\rm
Li}$=71$^{+18}_{-13}$\,m\AA, where the statistical uncertainties are
evaluated using a bootstrap resampling technique from the statistical
uncertainties in the HIRES spectrum.  However, the systematic uncertainties associated
with the continuum normalization are likely comparable to
the amplitude of this feature; we therefore treat the result of this
measurement as a $2\sigma$ upper limit: ${\rm EW}_{\rm Li}$$<$107\,m\AA.

@@

{\it Age: Summary}---The main indicators for the youth of
TIC~141146667 are {\it i)} that it is a complex periodic variable, and
{\it ii)} that it is 1.5 magnitudes brighter (four times more
luminous) than main sequence stars of the same color, while showing no
indicators for binarity.  The first item constitutes evidence for
youth since all known CPVs are $\lesssim$200\,Myr \cite{Bouma2024}.
The reference sequences for the Pleiades (112\,Myr, \cite{Dahm2015})
show a few stars of equal luminosity and the same temperature,
suggesting a photometric isochronal age upper limit
$\lesssim$150\,Myr.  The weak lithium absorption suggests an age of at
least 20\,Myr based on an empirical comparison using Gaia-ESO spectra,
or at least 35\,Myr based on the \cite{Feiden2016} isochrones.  These
considerations yield our adopted age range of 150\,Myr.
